maxent model avec xvalidation

### 2019-04-04 src/maxent_avec_xvalidation.Rmd
removal of auc testing as there is no testing at this stage
looped model selection

```{r train and test on 2003 inc AIC for reg val}
train2003 <- subset(janprback, year != "2003")
test2003 <-  subset(janprback, year == "2003")
train2003pa <- train2003$occurrence
test2003p <- subset(test2003, occurrence == "1")
test2003b <- subset(test2003, occurrence == "0")
train2003  <- subset(train2003, select = -c(year, occurrence))

mod <- maxent(x = train2003, p = train2003pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE"))
mod
ev <- evaluate(p = test2003p, a= test2003b, model = mod)
ev
threshold(ev)
```

# beta value (regularization) selection

mod12 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "replicatetype=crossvalidate", "replicates=5")) #add to samples because field study with biased data - see https://esajournals.onlinelibrary.wiley.com/doi/10.1890/10-1171.1. note this should be the default but keep here as a note

>>>>

```{r train and test on jan inc AIC for reg val}
modelmonth <- cbind(paste0("jan",unique(janprback$year)))
AIClist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
betavallist <- cbind(AIClist$regularization)

for (year in 1:nrow(modelmonth)){
  timeslice <- modelmonth[year] #testing timeslice
  yr <- str_sub(timeslice, -4) #the year being used for testing
  train <- subset(janprback, year != yr) #training data is all years but testing
  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
  test <-  subset(janprback, year == yr) #creates the testing dataset
  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling

  for (h in 1:nrow(betavallist)){
    betamval <-  betavallist[h]
    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))

    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code

    #count number of non-zero coefficients, excluding the final four rows
    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
    coefile <- head(coefile, -4)
    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
    nocoeff <- nrow(coefile2)
    bick <- sum(as.numeric(coefile2[, 2]))
    bic <- (bick*log(sum(trainpa == 1)) - 2*11)

    #no presences used in model (less those used for testing)
    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    notrainingpoints <- maxresults$X.Training.samples
    testauc <- maxresults$Test.AUC
    trainauc <- maxresults$Training.AUC

    #natural log
    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    samplepred$nat_log <- log(samplepred$Raw.prediction)
    sumsamplepred <- sum(samplepred$nat_log)

    #test auc scores and other lovely things you might want to include
    ev <- evaluate(p = testp, a= testb, model = mod)
    evauc <- ev@auc
    evcor <- ev@cor
    evpcor <- ev@pcor
    evnp <- ev@np
    evna <- ev@na
    evmaxtprtnr <- threshold(ev, 'spec_sens')
    evmaxkap <- threshold(ev, 'kappa')
    evno_omission <- threshold(ev, 'no_omission')
    evprevalence <- threshold(ev, 'prevalence')
    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')

    for (i in 1:nrow(AIClist)){
      if(AIClist$regularization[i] == betamval){
      AIClist$Train.AUC[i] <- trainauc
      AIClist$n[i] <- notrainingpoints
      AIClist$coefficients[i] <- nocoeff
      AIClist$sum.of.logs[i] <- sumsamplepred
      AIClist$Test.AUC[i] <- evauc
      AIClist$evcor[i] <- evcor
      AIClist$evpcor[i] <- evpcor
      AIClist$nopresence[i] <- evnp
      AIClist$nobackground[i] <- evna
      AIClist$maxsumsenspec[i] <- evmaxtprtnr
      AIClist$maxsumsenspec[i] <- evmaxkap
      AIClist$no_omission[i] <- evno_omission
      AIClist$equal_sens_spec[i] <- evequal_sens_spec
      AIClist$bic[i] <- bic #not sure this is correct...
      }
    }
  }

AIClist$AIC <- ((2*AIClist$coefficients)-(2*AIClist$sum.of.logs))
AIClist$AICc <- (-2*(AIClist$sum.of.logs)+((2*AIClist$coefficients)*(AIClist$n/(AIClist$n - AIClist$coefficients-1))))

write.csv(AIClist, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)

}

```

next step... select the models you want for each year and run those models again (primarily based on lowest AICc but check test AUC and other things too)

jan2003: b = 1
jan2004: b = 1
jan2005: b = 1.5
jan2006: b = 2.25

for (year in 1:nrow(modelmonth)){
  timeslice <- modelmonth[year] #testing timeslice
  yr <- str_sub(timeslice, -4) #the year being used for testing
  train <- subset(janprback, year != yr) #training data is all years but testing
  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
  test <-  subset(janprback, year == yr) #creates the testing dataset
  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling

  for (h in 1:nrow(betavallist)){
    betamval <-  betavallist[h]
    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))

    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code

    #count number of non-zero coefficients, excluding the final four rows
    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
    coefile <- head(coefile, -4)
    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
    nocoeff <- nrow(coefile2)
    bick <- sum(as.numeric(coefile2[, 2]))
    bic <- (bick*log(sum(trainpa == 1)) - 2*11)

    #no presences used in model (less those used for testing)
    maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    notrainingpoints <- maxresults$X.Training.samples
    testauc <- maxresults$Test.AUC
    trainauc <- maxresults$Training.AUC

    #natural log
    samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    samplepred$nat_log <- log(samplepred$Raw.prediction)
    sumsamplepred <- sum(samplepred$nat_log)

    #test auc scores and other lovely things you might want to include
    ev <- evaluate(p = testp, a= testb, model = mod)
    evauc <- ev@auc
    evcor <- ev@cor
    evpcor <- ev@pcor
    evnp <- ev@np
    evna <- ev@na
    evmaxtprtnr <- threshold(ev, 'spec_sens')
    evmaxkap <- threshold(ev, 'kappa')
    evno_omission <- threshold(ev, 'no_omission')
    evprevalence <- threshold(ev, 'prevalence')
    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')

    for (i in 1:nrow(AIClist)){
      if(AIClist$regularization[i] == betamval){
      AIClist$Train.AUC[i] <- trainauc
      AIClist$n[i] <- notrainingpoints
      AIClist$coefficients[i] <- nocoeff
      AIClist$sum.of.logs[i] <- sumsamplepred
      AIClist$Test.AUC[i] <- evauc
      AIClist$evcor[i] <- evcor
      AIClist$evpcor[i] <- evpcor
      AIClist$nopresence[i] <- evnp
      AIClist$nobackground[i] <- evna
      AIClist$maxsumsenspec[i] <- evmaxtprtnr
      AIClist$maxsumsenspec[i] <- evmaxkap
      AIClist$no_omission[i] <- evno_omission
      AIClist$equal_sens_spec[i] <- evequal_sens_spec
      AIClist$bic[i] <- bic #not sure this is correct...
      }
    }
  }

AIClist$AIC <- ((2*AIClist$coefficients)-(2*AIClist$sum.of.logs))
AIClist$AICc <- (-2*(AIClist$sum.of.logs)+((2*AIClist$coefficients)*(AIClist$n/(AIClist$n - AIClist$coefficients-1))))

write.csv(AIClist, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)

}

### 2019-04-04 src/stack_test.Rmd
looped model selection

try changing some of the settings here - paerticularly you want aic
```{r}
VariableSelection(maxent="C:/Users/bunny/Downloads/maxent/maxent.jar", outdir=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/output"), gridfolder=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/testgrid"), occurrencelocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/presence_red.csv"), backgroundlocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/background_red.csv"), additionalargs="nohinge", contributionthreshold=5, correlationthreshold=0.9, betamultiplier=seq(0.25, 2, 0.25))
```
no you cant - you need to use the grids...

>>>>

try changing some of the settings here - particularly you want aic
```{r}
VariableSelection(maxent="C:/Users/bunny/Downloads/maxent/maxent.jar", outdir=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/output"), gridfolder=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/testgrid"), occurrencelocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/presence_red.csv"), backgroundlocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/background_red.csv"), additionalargs="nohinge", contributionthreshold=5, correlationthreshold=0.9, betamultiplier=seq(0.25, 2, 0.25))
```
no you cant - you need to use the grids...
